The Concurrent Calculi Formalisation Benchmark

### 2023-10-12 paper/paper.tex

We have begun collecting solutions to our challenge problems on our website:
%
\begin{center}
  \url{https://concurrentbenchmark.github.io/}
\end{center}
%
In the longer term, we expect to use the website for promoting best practices,
tutorials and guidelines derived from solutions to our challenge problems.
%The interested reader can find more information about the current status at:
We encourage anyone interested to try the challenge problems using their
favourite tools and techniques, and to send us their solutions.

\paragraph{Outline of this paper.}
In \cref{sec:challenge-problems} we describe in more detail the three key
aspects of our challenge problems.  We discuss our design choices in
\cref{sec:design-discussion}, and possible extensions and further topics in
\cref{sec:going-beyond}.  We conclude in \cref{sec:conclusion}.
The full challenge problems, including proofs, are available in \cref{app:challenges}.

\section{The challenge problems}\label{sec:challenge-problems}
In this section we briefly describe the three aspects of our challenge problems, referring to the appendix for the full text of each one.

@@

% We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types, psi-calculi, or encodings between different calculi are not directly covered.  Still, these (and other) proof mechanisation applications will need the basic techniques that our challenge problems exercise:
as per our design goal \ref{item:goal-reusability},
our problems are drawn from the basic metatheory of process calculi (instead
of requiring the development of new theory), and focus on well-known theorems
and results that involve interesting proof techniques that may be reused in further work.

To achieve our design goal \ref{item:goal-comperison-accessibility},
we have designed our challenge problems to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.  The process calculus used in each challenge
focuses on the features we want to exercise, and we have omitted all constructs
(such as choices) that would complicate the mechanisation without bringing tangible
benefits and insights.  (Adding the omitted constructs is of course possible, but we leave
such extensions as a possibility for a future benchmark challenge.)

@@

We have chosen a simple notion of well-formedness for our challenge problem,
and do not for instance consider deadlocked processes. While more
sophisticated notions of well-formedness are also interesting, the
proofs of these properties are more complicated and would thus make
the challenge problem more difficult and less focused on the linear
aspects of the system.

@@

\subsection{Evaluation criteria}
Solutions to our challenge problems should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges need to present more elaborate argumentation for their adequacy.

>>>>

We have begun collecting solutions to our challenges on our website:
%
\begin{center}
  \url{https://concurrentbenchmark.github.io/}
\end{center}
%
In the longer term, we expect to use the website for promoting best practices,
tutorials and guidelines derived from solutions to our challenges.
%The interested reader can find more information about the current status at:
We encourage anyone interested to try the challenges using their
favourite tools and techniques, and to send us their solutions.

\paragraph{Outline of this paper.}
In \cref{sec:challenge-problems} we describe in more detail the three key
aspects of our challenge problems.  We discuss our design choices in
\cref{sec:design-discussion}, and possible extensions and further topics in
\cref{sec:going-beyond}.  We conclude in \cref{sec:conclusion}.
The full challenges, including proofs, are available in \cref{app:challenges}.

\section{The challenges}\label{sec:challenge-problems}
In this section we briefly describe the three aspects of our challenges, referring to the appendix for the full text of each one.

@@

% We have identified three key aspects fundamental to mechanisations of concurrent calculi and designed our challenge problems to exercise these.
Like the previous POPLMark and POPLMark Reloaded challenges, our challenge is not meant to be comprehensive: applications such as multiparty session types, choreographies, conversation types, psi-calculi, or encodings between different calculi are not directly covered.  Still, these (and other) proof mechanisation applications will need the basic techniques that our challenges exercise:
as per our design goal \ref{item:goal-reusability},
our problems are drawn from the basic metatheory of process calculi (instead
of requiring the development of new theory), and focus on well-known theorems
and results that involve interesting proof techniques that may be reused in further work.

To achieve our design goal \ref{item:goal-comperison-accessibility},
we have designed our challenges to exercise the three aspects (linearity, scope extrusion, coinduction) independently, and they can be solved individually and in any order.
We have designed each problem to be small and easy to understand with basic knowledge of textbook concurrency theory, process calculi and type theory.  The process calculus used in each challenge
focuses on the features we want to exercise, and we have omitted all constructs
(such as choices) that would complicate the mechanisation without bringing tangible
benefits and insights.  (Adding the omitted constructs is of course possible, but we leave
such extensions as a possibility for a future benchmark challenge.)

@@

We have chosen a simple notion of well-formedness for our challenge,
and do not for instance consider deadlocked processes. While more
sophisticated notions of well-formedness are also interesting, the
proofs of these properties are more complicated and would thus make
the challenge more difficult and less focused on the linear
aspects of the system.

@@

\subsection{Evaluation criteria}
Solutions to our challenges should be compared on three measures: the mechanisation overhead, this being the amount of manually written infrastructure and setup needed to express the definitions in the mechanisation; the adequacy of the formal statements in the mechanisation, i.e.\ whether the proven theorems are easily recognisable as the theorems from the challenge; and the cost of entry for the tools and techniques employed, i.e.\ the difficulty of learning to use the techniques.
Solutions to our challenges do not need to strictly follow the definitions and lemmas set out in the problem text, but solutions which deviate from the original challenges need to present more elaborate argumentation for their adequacy.
